Comparative Analysis of Approximation Algorithms for Solving TSP

### 2018-04-13 tex/paper.tex
Fix URLs out of margin
Remove the trademark from Octave
According to GNU recommendations, no trademark symbol should be included

\usepackage{bm}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{url}
\usepackage{multirow}

\begin{document}	
	\twocolumn[
	\begin{@twocolumnfalse}
		\author{Mohamed Moanis Ali, Michael Sedrak \\  mohamed.moanis.ali@gmail.com, m.sedrak@live.com}
		\title{Solving TSP With Genetic Algorithms}
		\maketitle
		\begin{abstract}
			The traveling salesman problem is considered one of hardest problems in computer science with no polynomial runtime algorithm to solve it so far. Nevertheless, TSP is an active research point with more algorithms being developed to either approximately solve the original problem, or to speed it up so larger datasets can be solved. In this work we present a comparative study using four algorithms to solve TSP -genetic algorithm, simulated annealing, nearest neighbor and ant colony optimization to analyze the runtime performance and the correctness of their solution.  The results of the experiments on the Uruguay dataset -using both {Matlab\texttrademark}  and {Octave\texttrademark}- showed that the nearest neighbor algorithm provided the fastest runtime and a relatively correct solution with $18\%$ from the known optimal solution, while the ant colony optimization proved to be the least correct among the studied algorithms. The study also found a huge performance difference between {Matlab\texttrademark}  and {Octave\texttrademark} software.
		\end{abstract}

@@

	\section{Introduction}
	The traveling salesman problem (TSP) is a classical combinatorial optimization problem with a long history -dating back to 1832, and countless number of real life applications\cite{applegate07}\cite{punnen07}. It was first mentioned in a handbook for traveling salesmen in Germany and Switzerland to be mathematically formulated later in 1930. Though easy to describe, TSP is a NP-hard problem; that's it, there is no polynomial time algorithm to solve it. This limits the use of exact methods which have a time complexity of $O(n!)$ to a small number of cities.
	The problem is to find the shortest route between a given set of cities' locations where each city must be visited only once and the trip must end by returning back to the starting city. TSP can be modeled as an undirected weighted graph, such that cities are the graph's vertices and paths between cities are the graph's edges. The distance between two given cities in both direction can be the same which is called symmetric TSP and can differ in case of asymmetric variant. In this work, we will only consider the symmetric TSP.
	The optimization problem to solve TSP can be formulated as a linear programming problem as follows\cite{papadimitriou98}:
	\begin{equation} \label{eq:TSPLinearProgramming}
	min \sum_{i=1}^{n} \sum_{j=1, j\neq i}^{n} {c_i}_j {x_i}_j
	\end{equation}
	where ${c_i}_j$ is the cost between cities $i$ and $j$ and ${x_i}_j$ is given by Eq.(\ref{eq:TSPX})
	\begin{equation} \label{eq:TSPX}
	{x_i}_j = 
	\begin{cases}
	1,& \text{path goes from i to j}\\
	0,& \text{otherwise}
	\end{cases}
	\end{equation}
	The complexity of TSP, makes room for approximate algorithms to arise\cite{Brucal17}. Such algorithms can reach near optimal solution in a reasonable time as will be seen in this work. Algorithms such as variable neighborhood search\cite{Thanh15}, genetic algorithm\cite{Wang17}, monte-carlo trees search\cite{Perez14}, adaptive bee colony\cite{Rekaby13}, meta-learning\cite{Kanda11}, machine learning\cite{Pihera14}, simulated annealing\cite{Kerrache14}, ant colony optimization\cite{Swiatnicki15}, harmony search\cite{Tongchan17}, biography based approximation\cite{Wu17} and variations of neural networks\cite{Gao10}\cite{Mueller15} have addressed the TSP. In this work we will focus on four of such algorithms, namely: genetic algorithm\cite{Kirk14}\cite{mathwork}, simulated annealing\cite{Jang02}, nearest neighbor search\cite{Jevtic14} and ant colony optimization\cite{Ibrahim15}. This survey will compare {Matlab\texttrademark}  implementation of the four algorithms in terms of runtime performance and how close they are to the optimal solution.

@@

	\section{Conclusion}
	In this study, we have provided a comparative analysis between the four algorithms -GA, SA, NN, and ACO- for solving the TSP using free {Matlab\texttrademark} implementations. The analysis was focused on measuring the runtime performance and how close is the solution from the known optimal solution of the the Uruguay dataset of 734 unique points. Furthermore, the experiments were run on both {Matlab\texttrademark} and {Octave\texttrademark} software to highlight the differences between the two in terms of performance.

@@

	Additionally, the difference between the runtime performance of the four algorithms when run on both {Matlab\texttrademark} and {Octave\texttrademark} was found to be very huge; for example in the case of GA, {Matlab\texttrademark} was 380x faster. This can draw a conclusion that the current state of {Matlab\texttrademark} software far exceeds it's free open source alternative {Octave\texttrademark}.

>>>>

\usepackage{bm}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[hyphens]{url}
\usepackage{multirow}

\begin{document}	
	\twocolumn[
	\begin{@twocolumnfalse}
		\author{Mohamed Moanis Ali, Michael Sedrak \\  mohamed.moanis.ali@gmail.com, m.sedrak@live.com}
		\title{Comparative Analysis of Approximation Algorithms for Solving TSP}
		\maketitle
		\begin{abstract}
			The traveling salesman problem is considered one of hardest problems in computer science with no polynomial runtime algorithm to solve it so far. Nevertheless, TSP is an active research point with more algorithms being developed to either approximately solve the original problem, or to speed it up so larger datasets can be solved. In this work we present a comparative study using four algorithms to solve TSP -genetic algorithm, simulated annealing, nearest neighbor and ant colony optimization to analyze the runtime performance and the correctness of their solution.  The results of the experiments on the Uruguay dataset -using both {Matlab\texttrademark}  and Octave- showed that the nearest neighbor algorithm provided the fastest runtime and a relatively correct solution with $18\%$ from the known optimal solution, while the ant colony optimization proved to be the least correct among the studied algorithms. The study also found a huge performance difference between {Matlab\texttrademark}  and Octave software.
		\end{abstract}

@@

	\section{Introduction}
	The traveling salesman problem (TSP) is a classical combinatorial optimization problem with a long history -dating back to 1832, and countless number of real life applications\cite{applegate07}\cite{punnen07}. It was first mentioned in a handbook for traveling salesmen in Germany and Switzerland to be mathematically formulated later in 1930. Though easy to describe, TSP is a NP-hard problem; that's it, there is no polynomial time algorithm to solve it. This limits the use of exact methods which have a time complexity of $O(n!)$ to a small number of cities.

	The problem is to find the shortest route between a given set of cities' locations where each city must be visited only once and the trip must end by returning back to the starting city. TSP can be modeled as an undirected weighted graph, such that cities are the graph's vertices and paths between cities are the graph's edges. The distance between two given cities in both direction can be the same which is called symmetric TSP and can differ in case of asymmetric variant. In this work, we will only consider the symmetric TSP.

	The optimization problem to solve TSP can be formulated as a linear programming problem as follows\cite{papadimitriou98}:
	\begin{equation} \label{eq:TSPLinearProgramming}
	min \sum_{i=1}^{n} \sum_{j=1, j\neq i}^{n} {c_i}_j {x_i}_j
	\end{equation}
	where ${c_i}_j$ is the cost between cities $i$ and $j$ and ${x_i}_j$ is given by Eq.(\ref{eq:TSPX})
	\begin{equation} \label{eq:TSPX}
	{x_i}_j = 
	\begin{cases}
	1,& \text{path goes from i to j}\\
	0,& \text{otherwise}
	\end{cases}
	\end{equation}
	The complexity of TSP, makes room for approximate algorithms to arise\cite{Brucal17}. Such algorithms can reach near optimal solution in a reasonable time as seen in this work. Algorithms such as variable neighborhood search\cite{Thanh15}, genetic algorithm\cite{Wang17}, monte-carlo trees search\cite{Perez14}, adaptive bee colony\cite{Rekaby13}, meta-learning\cite{Kanda11}, machine learning\cite{Pihera14}, simulated annealing\cite{Kerrache14}, ant colony optimization\cite{Swiatnicki15}, harmony search\cite{Tongchan17}, biography based approximation\cite{Wu17} and variations of neural networks\cite{Gao10}\cite{Mueller15} have addressed the TSP. In this work we will focus on four of such algorithms, namely: genetic algorithm\cite{Kirk14}\cite{mathwork}, simulated annealing\cite{Jang02}, nearest neighbor search\cite{Jevtic14} and ant colony optimization\cite{Ibrahim15}. This survey will compare {Matlab\texttrademark}  implementation of the four algorithms in terms of runtime performance and how close they are to the optimal solution.

@@

	\section{Conclusion}
	In this study, we have provided a comparative analysis between the four algorithms -GA, SA, NN, and ACO- for solving the TSP using free {Matlab\texttrademark} implementations. The analysis was focused on measuring the runtime performance and how close is the solution from the known optimal solution of the the Uruguay dataset of 734 unique points. Furthermore, the experiments were run on both {Matlab\texttrademark} and Octave software to highlight the differences between the two in terms of performance.

@@

	Additionally, the difference between the runtime performance of the four algorithms when run on both {Matlab\texttrademark} and Octave was found to be very huge; for example in the case of GA, {Matlab\texttrademark} was 380x faster. This can draw a conclusion that the current state of {Matlab\texttrademark} software far exceeds it's free open source alternative Octave.
